---
title: "Assignment_3"
author: "Gloria"
date: "2022-10-17"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
```{r}
library(readr)
UniversalBank <- read_csv("C:/Users/idast/OneDrive/Desktop/UniversalBank.csv")

```
```{r}
#loading Libraries
library(readr)
library(caret)
library(ISLR)
library(class)

```

```{r}
View(UniversalBank)
DF= UniversalBank
DF$Online_category='Not-Active'
DF$Online_category[DF$Online>0]= 'Active'
DF$Online_category=as.factor(DF$Online_category)
DF$CreditCard=as.factor(DF$CreditCard )
DF$PersonalLoan=as.factor(DF$PersonalLoan)

```

```{r}
summary(DF)
```
#TASK A. 
#Partition of the data is segregated into 60 and 40 with a pivot table for training data with Online as column variable and CreditCard as row and loan as secondary variable.

#Using the xtabs and ftable functions.

```{r}
set.seed(64060)
Train_Index = createDataPartition(DF$PersonalLoan,p=0.6, list=FALSE)
Train.df=DF[Train_Index,]
```


```{r}
mytable <- xtabs(~ Online_category+CreditCard+PersonalLoan, data=Train.df)
ftable(mytable)
```
#Task B
#The probability of loan acceptance (Loan = 1) conditionally on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)] is 59/(59+479) = 0.10966 OR 11%


#Task C 
#Creating Pivot tables for training data with one of CreditCard (rows) as a function of PersonalLoan(columns) and another  of Online category(rows) with PersonalLoan(columns).

```{r}

table(Creditcard =Train.df$CreditCard, PersonalLoan =Train.df$PersonalLoan)


```



```{r}
table(Online_category =Train.df$Online_category, PersonalLoan =Train.df$PersonalLoan)
```

#Task-D
#i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors) Ans. 93/(93+195) = 0.323 in otherwords 32.3%

#ii. P(Online = 1 | Loan = 1) Ans.179/(179+109)= 62.51%

#iii.P(Loan = 1) (the proportion of loan acceptors) Ans. 195+93/(195+93+1924+788)= 0.096 or in otherwords 9.6%

#iv. P(CC = 1 | Loan = 0) Ans.788/(1924+788)= 0.2905 or 29.05%

#v. P(Online = 1 | Loan = 0) Ans.1631/(1631+1081)=0.6014 or 60.14%

#vi. P(Loan = 0) Ans.1924+788/(1924+788+195+93)=0.904 or 90.4%



#Task-E
#Using the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online
= 1).
#formula P(Y/x1. . . Xn) = (P(X1. . . . . . Xn/Y)/P(X1,. . . .Xn)) X1=Online X2=Creditcard Y=PersonalLoan

P(creditcard) =0.60333 P(Online) =0.2936662 = = 0.6251*0.323*0.096/0.60333*0.2936662=0.105598 or 10.55%



#TASK-F
#Both normal method (10.96%) and Naive Bayes method values (10.55%) are very similar.Both are similar, with the difference being Naive Bayes assumes attributes are independent of each other making it less accurate. 



#TASK-G 
#Running Naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}
library(e1071)
nb.model<-naiveBayes (PersonalLoan~Online_category+CreditCard, data=Train.df)

To_Predict=data.frame(CreditCard ='1',Online_category ='1')
predict(nb.model,To_Predict,type='raw')
```










