---
title: "ML2"
author: "Gloria Stephen"
date: "2022-10-02"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(caret)
UniversalBank <- read_csv("C:/Users/idast/Downloads/UniversalBank.csv")
UniversalBank$Personal.Loan =as.factor(UniversalBank$Personal.Loan)
data1<-UniversalBank

summary(UniversalBank)

View(data1)
```
##removing unwanted data from dataset
```{r}
data2<-data1[,-c(1,5)]
View(data2)
```

#converting personal loan to factors
```{r}
#data2$Personal.Loan =as.factor(data2$Personal.Loan)

View(data2)
```

#creating the dummy variables 
```{r}
Education_1 <- ifelse(data2$Education == 1, 1,0)
Education_2 <- ifelse(data2$Education == 2, 1,0)
Education_3 <- ifelse(data2$Education == 3, 1,0)
```


```{r}

data3<-data.frame(Age=data2$Age,Experience=data2$Experience,Income=data2$Income,Family=data2$Family,CCAvg=data2$CCAvg, Education_1=Education_1,Education_2=Education_2,Education_3=Education_3,Personal.Loan=data2$Personal.Loan,Mortgage=data2$Mortgage,Securities.Account=data2$Securities.Account,CD.Account=data2$CD.Account,Online=data2$Online,CreditCard=data2$CreditCard)

View(data3)
```

##creating a partition of 60:40 
```{r}
Train_Index = createDataPartition(data3$Personal.Loan,p=0.6, list = FALSE)
Train.df =data3[Train_Index,]
Validation.df=data3[-Train_Index,]

nrow(Train.df)

nrow(Validation.df)
```

#normalization of the data

```{r}

Norm_model <- preProcess(Train.df[,-(6:9)], method = c("center", "scale"))

training_norm<-predict(Norm_model,Train.df)

validation_norm<-predict(Norm_model,Validation.df)


```
#test data set

```{r}
Test<-data.frame(Age=40,Experience=10,Income=84,Family=2,CCAvg=2,Education_1=0,Education_2=1,Education_3=0,Mortgage=0,Securities.Account=0,CD.Account=0,Online=1,CreditCard=1)


View(Test)

test_norm<-predict(Norm_model,Test)

```
#knn algorithm in dataset
```{r}


library(class)

pred_train<-training_norm[,-9]
label_train<-training_norm[,9]

pred_valid<-validation_norm[,-9]
label_valid<-validation_norm[,9]



View(training_norm)

Model<-knn(pred_train,test_norm,cl=label_train,k=1)

Model


#since value of k=0, the customer will not accept the loan offer


```
#Finding the best value for k

```{r}

set.seed(541)

searchGrid <- expand.grid(k=seq(1:30))

predicted <- train(Personal.Loan~.,training_norm,method="knn",tuneGrid=searchGrid)

predicted

best_k <- predicted$bestTune[[1]]

#this saves the best value for k

```

#the confusion matrix

```{r}

model<-predict(predicted,validation_norm[,-9])

confusionMatrix(model,label_valid)


```

#knn for new customer

```{r}
Prediction_new<-knn(pred_train,test_norm,cl=label_train,k=best_k)

Prediction_new

#k=0, new customer will not accept the loan offer

```
#5#splitting data to 50:30:20 ratio

set.seed(887)


part1<- createDataPartition(data3$Personal.Loan,p=.5,list=FALSE,times=1)

part2<- createDataPartition(data3$Personal.Loan,p=.3,list=FALSE,times=1)

part3<- createDataPartition(data3$Personal.Loan,p=.2,list=FALSE,times=1)

new_train  <- data3[part1, ]

new_valid  <- data3[part2, ]

new_test <- data3[part3, ]


#normalization of the data

normalized <- preProcess(Train.df[,-(6:9)], method=c("center","scale"))

#training data 

normalized_train <- predict(normalized,new_train)

#the validation data 

normalized_valid<-predict(normalized,new_valid)

#the test data created

normalized_test<-predict(normalized,new_test)

#running knn for train,validation and test data

pred_train1=normalized_train[,-9]

label_train1=normalized_train[,9]

pred_valid1=normalized_valid[,-9]

label_valid1=normalized_valid[,9]

pred_test1=normalized_test[,-9]

label_test1=normalized_test[,9]

training_prediction<-knn(pred_train1,pred_train1,cl=label_train1,k=best_k)

validation_prediction<-knn(pred_train1,pred_valid1,cl=label_train1,k=best_k)

test_prediction<-knn(pred_train1,pred_test1,cl=label_train1,k=best_k)

#confusion matrix of training data

confusionMatrix(pred_train1,label_train1)


#confusion matrix of validation data

confusionMatrix(pred_valid1,label_valid1)


#confusion matrix of test data


confusionMatrix(pred_test1,label_test1)


